---
title: "Examples of Nonprofits Within Tax Exempt Purpose Categories"
format:
  html:
    html-table-processing: none
    code-fold: true
    toc: true
    toc-location: left
    embed-resources: true
number-sections: false
execute:
  warning: false
  message: false
  echo: false
---


### New Primary Tax Exempt Purpose Categories

- **BUSINESS AND LABOR**: business and labor associations 
- **CHARITY**: public charities 
- **CIVIC/ADVOCACY**: civic engagement, advocacy, and politically-engaged organizations
- **COOPERATIVES**: joint assurances and protections from risk; wealth building and asset building.
- **DEFUNCT**: category no longer supported by the IRS 
- **INSURANCE**: protection from risk (primarily individuals, also orgs)
- **MEMBER ASSOCIATIONS**: groups that operate for the benefit of members (other than cooperatives and pensions) 
- **PENSION**: wealth building orgs, income smoothing, savings mechanisms
- **PRIVATE FOUNDATION**: private foundations 
- **SPECIAL PURPOSE CORP**: tax-exempt corporations with a narrowly-defined set of allowable operations 



<br>
<hr>
<br>




```{r, echo=F}
library( dplyr )
library( stringr )
library( pander )
library( knitr )
library( data.table )
library( DT )
library( kableExtra )

# setwd( "C:/Users/jdlec/Dropbox/00 - URBAN/taxonomies" )
```




```{r}
new.order <- 
  c("EIN", "EIN2",
    "ORG_NAME_CURRENT","ORG_NAME_SEC",
    "ORG_FISCAL_YEAR", "ORG_PERS_ICO", 
    "ORG_FISCAL_PERIOD",
    "ORG_RULING_DATE", "ORG_RULING_YEAR", 
    ### CORE COUNTS
    "ORG_YEAR_FIRST", "ORG_YEAR_LAST", 
    "ORG_YEAR_COUNT", 
    "NTEE_IRS", "NTEE_NCCS", "NTEEV2", 
    "NCCS_LEVEL_1", 
    "NCCS_LEVEL_2", 
    "NCCS_LEVEL_3",
    ### BMF
    "BMF_SUBSECTION_CODE", "BMF_STATUS_CODE", 
    "BMF_PF_FILING_REQ_CODE", "BMF_ORGANIZATION_CODE", 
    "BMF_INCOME_CODE", "BMF_GROUP_EXEMPT_NUM", 
    "BMF_FOUNDATION_CODE", "BMF_FILING_REQ_CODE", 
    "BMF_DEDUCTIBILITY_CODE", "BMF_CLASSIFICATION_CODE", 
    "BMF_ASSET_CODE", "BMF_AFFILIATION_CODE",
    ### FISCAL FIELDS 
    "F990_TOTAL_REVENUE_RECENT", 
    "F990_TOTAL_INCOME_RECENT", 
    "F990_TOTAL_ASSETS_RECENT", 
    "F990_ORG_ADDR_CITY", 
    "F990_ORG_ADDR_STATE", 
    "F990_ORG_ADDR_ZIP", 
    "F990_ORG_ADDR_STREET",
    ### GEOCODE
    "ORG_ADDR_FULL", "ORG_ADDR_MATCH", 
    "LATITUDE", "LONGITUDE", 
    "GEOCODER_SCORE", "GEOCODER_MATCH", 
    ### CENSUS
    "CENSUS_CBSA_FIPS", 
    "CENSUS_CBSA_NAME", 
    "CENSUS_BLOCK_FIPS", 
    "CENSUS_URBAN_AREA", 
    "CENSUS_STATE_ABBR", 
    "CENSUS_COUNTY_NAME"  )
```


```{r, cache=TRUE}
# setwd( "C:/Users/jdlec/Dropbox/00 - URBAN/taxonomies" )
d <- data.table::fread( "BMF_UNIFIED_V1.1.csv" ) %>% as.data.frame()
d <- d[ new.order ]

ss <- read.csv("sscc.csv")
ss2 <- ss[c("SSCC","SSTYPE","SSTYPE2","CCTYPE","SSCCTYPE1","SSCCTYPE2")]

d.types <- read.csv("TAX-EXEMPT-TYPES-V2.csv")

keep <- 
c("bmf_code", 
  "irs_tax_code", 
  "v501c_label", 
  "v501c_desc", 
  "v501c_desc2", 
  "exempt_type_stakeholder", 
  "exempt_type_primary", 
  "exempt_type_secondary", 
  "govt_incorp", 
  "req_990", 
  "option_990EZ", 
  "donations_deduct", 
  "political_limits", 
  "member_types", 
  "affiliation_status")

d.types2 <- d.types[keep]

# DROP PRIVATE FOUNDATIONS BECAUSE IT DUPLICATES 501c3 ROWS
d.types2 <- dplyr::filter( d.types2, v501c_desc2 != "501(c)(3) - Private Foundations" )
```



```{r}
dd <- 
  d.types2 %>% 
  select( exempt_type_primary, 
          irs_tax_code, v501c_label, 
          exempt_type_stakeholder )

names(dd) <- toupper( names(dd) )

dd %>% 
  kable( escape=FALSE, format = "html" ) %>%  
  kable_paper()
```


```{r}
## Classification Codes are 4-digits

d$CC1 <- substr( d$BMF_CLASSIFICATION_CODE, 1, 1 )
d$CC2 <- substr( d$BMF_CLASSIFICATION_CODE, 2, 2 )
d$CC3 <- substr( d$BMF_CLASSIFICATION_CODE, 3, 3 )
d$CC4 <- substr( d$BMF_CLASSIFICATION_CODE, 4, 4 )

d$SS   <- stringr::str_pad( d$BMF_SUBSECTION_CODE, 2, pad = "0" )
d$SS   <- paste0( "c", d$SS )
d$SSCC <- paste0( d$SS, "-", d$CC1 )
```


```{r}
ubmf <- merge( d, d.types2, by.x="BMF_SUBSECTION_CODE", by.y="bmf_code", all.x=T )
ubmf <- merge( ubmf, ss2, by="SSCC" )
```




```{r}
ubmf$SSCCTYPE2 <- paste0( "501", ubmf$SSCCTYPE2 )

keep.ubmf <- 
c("irs_tax_code","SSTYPE2",
  "ORG_NAME_CURRENT","ORG_NAME_SEC", 
  
  "CCTYPE","SSTYPE",
  "exempt_type_primary","exempt_type_secondary", 
  "govt_incorp", "req_990", 
  "option_990EZ", "donations_deduct", 
  "political_limits", "member_types", 
  "affiliation_status","SS",
  "SSCCTYPE2",
  "ORG_FISCAL_YEAR", "ORG_RULING_YEAR", 
  "NTEEV2","NCCS_LEVEL_2",
  "F990_TOTAL_INCOME_RECENT", "F990_TOTAL_ASSETS_RECENT",
  "F990_ORG_ADDR_CITY")

u2 <- ubmf[ keep.ubmf ]
u2 <- dplyr::filter( u2, SS != "c00" & CCTYPE != "not defined in the bmf dictionary" )

primary.exempt.types <- unique( u2$exempt_type_primary ) |> sort()

na.fis <- is.na( u2$ORG_FISCAL_YEAR )
na.inc <- is.na( u2$F990_TOTAL_INCOME_RECENT )
na.ass <- is.na( u2$F990_TOTAL_ASSETS_RECENT )
u2 <- dplyr::filter( u2, ! ( na.fis | na.inc | na.ass ) )
```


### Num of Nonprofits by Primary Tax Type in the BMF


```{r}
t <- 
  u2 %>%
  group_by( exempt_type_primary ) %>%
  summarize( N=n() ) %>%
  ungroup() 

t %>% 
  kable( escape=FALSE, format = "html" ) %>%  
  kable_paper()
```



```{r, results="asis"}
u2$SSTYPE2 <- paste0( "501", u2$SSTYPE2 )

for( i in primary.exempt.types )
{
  
  bmf.sub <- 
    u2 %>% 
    dplyr::filter( exempt_type_primary == i ) %>%
    dplyr::select( - c(CCTYPE,SSTYPE,SS,irs_tax_code) )
    
  cat( paste0( "## ", i, "  \n\n" ),  sep = "" )
    
  samp <- dplyr::sample_n( bmf.sub, min(nrow(bmf.sub),25) )
  names(samp) <- toupper( names(samp) )
  
  cat( pander::pander( samp ),  sep = "" ) 
    
  cat( paste0( "\n<br><br><br><br>  \n\n" ),  sep = "" )
}
```



```{=html}
<style>

h2 {
  margin-top: 80px;
  color: steelblue;
}

.table th { color: darkorange !important; }

.table td { font-family: monospace, monospace !important; }


</style>
```