---
title: "Exploring Tax Exempt Status "
output: html_notebook
---


## Import data 

```{r}
library(tidyverse)
library(DT)

### Master BMF
bmf.nccs <- read_csv("https://nccs-data.urban.org/dl.php?f=bmf/2022/bmf.bm2208.csv", 
                col_types = cols(col_factor(NULL)))

bmf.irs <- read_csv("https://nccsdata.s3.amazonaws.com/raw/bmf/2023-08-BMF.csv")

#drop columns we aren't using 
bmf.irs <- bmf %>% select(-c(ADDRESS, CITY, STATE ))

## Making classification codes - to look at with subsection codes
bmf.irs <- bmf.irs %>% 
  mutate(class1 = grepl("1", bmf$CLASSIFICATION),
          class2 = grepl("2", bmf$CLASSIFICATION),
          class3 = grepl("3", bmf$CLASSIFICATION),
          class4 = grepl("4", bmf$CLASSIFICATION),
          class5 = grepl("5", bmf$CLASSIFICATION),
          class6 = grepl("6", bmf$CLASSIFICATION),
          class7 = grepl("7", bmf$CLASSIFICATION),
          class8 = grepl("8", bmf$CLASSIFICATION))


### Making activity codes 
bmf.irs <- bmf.irs %>% 
  mutate(ACTIVITY = as.numeric(ACTIVITY) ) %>%
  mutate(activity1 = substr(ACTIVITY, 1, 3),
         activity2 = substr(ACTIVITY, 4, 6),
         activity3 = substr(ACTIVITY, 7, 9)) %>%
  mutate(activity1.group = case_when(
    activity1 < 30 ~ "religious activities",
    activity1 < 60 ~ "Schools, Colleges and Related Activities", 
    activity1 < 120 ~ "Cultural, Historical of Other Educational Activities",
    activity1 < 150 ~ "Other Instructions and Training Activities", 
    activity1 < 180 ~ "Health Services and Related Activities",
    activity1 < 200 ~ "Scientific Research Activities",
    activity1 < 230 ~  "Business and Professional Organizations", 
    activity1 < 250 ~ "Farming and Related Activities",
    activity1 < 260 ~ "Mutual Organizations",
    activity1 < 280 ~ "Employee of Membership Benefit Organizations",
    activity1 < 320 ~ "Sports, Athletic Recreational and Social Activities",
    activity1 < 350 ~ "Youth Activities",
    activity1 < 380 ~ "Conservation, Environmental and Beautification Activities",
    activity1 < 400 ~ "Housing Activities",
    activity1 < 430 ~ "Inner City or Community Activities",
    activity1 < 450 ~ "Civil Rights Activities",
    activity1 < 470 ~ "Litigation and Legal Aid Activities", 
    activity1 < 510 ~ "Legislative and Political Activities",
    activity1 < 560 ~ "Advocacy",
    activity1 < 600 ~ "Other Activities Directed to Individuals",
    activity1 < 605 ~ "Activities Purposes and Activities",
    activity1 < 450 ~ "Other Purposes and Activities",
    TRUE ~ "NONE" )) %>% 
  mutate(activity2.group = case_when(
    activity2 < 30 ~ "religious activities",
    activity2 < 60 ~ "Schools, Colleges and Related Activities", 
    activity2 < 120 ~ "Cultural, Historical of Other Educational Activities",
    activity2 < 150 ~ "Other Instructions and Training Activities", 
    activity2 < 180 ~ "Health Services and Related Activities",
    activity2 < 200 ~ "Scientific Research Activities",
    activity2 < 230 ~  "Business and Professional Organizations", 
    activity2 < 250 ~ "Farming and Related Activities",
    activity2 < 260 ~ "Mutual Organizations",
    activity2 < 280 ~ "Employee of Membership Benefit Organizations",
    activity2 < 320 ~ "Sports, Athletic Recreational and Social Activities",
    activity2 < 350 ~ "Youth Activities",
    activity2 < 380 ~ "Conservation, Environmental and Beautification Activities",
    activity2 < 400 ~ "Housing Activities",
    activity2 < 430 ~ "Inner City or Community Activities",
    activity2 < 450 ~ "Civil Rights Activities",
    activity2 < 470 ~ "Litigation and Legal Aid Activities", 
    activity2 < 510 ~ "Legislative and Political Activities",
    activity2 < 560 ~ "Advocacy",
    activity2 < 600 ~ "Other Activities Directed to Individuals",
    activity2 < 605 ~ "Activities Purposes and Activities",
    activity2 < 450 ~ "Other Purposes and Activities",
    TRUE ~ "NONE" )) %>% 
  mutate(activity3.group = case_when(
    activity3 < 30 ~ "religious activities",
    activity3 < 60 ~ "Schools, Colleges and Related Activities", 
    activity3 < 120 ~ "Cultural, Historical of Other Educational Activities",
    activity3 < 150 ~ "Other Instructions and Training Activities", 
    activity3 < 180 ~ "Health Services and Related Activities",
    activity3 < 200 ~ "Scientific Research Activities",
    activity3 < 230 ~  "Business and Professional Organizations", 
    activity3 < 250 ~ "Farming and Related Activities",
    activity3 < 260 ~ "Mutual Organizations",
    activity3 < 280 ~ "Employee of Membership Benefit Organizations",
    activity3 < 320 ~ "Sports, Athletic Recreational and Social Activities",
    activity3 < 350 ~ "Youth Activities",
    activity3 < 380 ~ "Conservation, Environmental and Beautification Activities",
    activity3 < 400 ~ "Housing Activities",
    activity3 < 430 ~ "Inner City or Community Activities",
    activity3 < 450 ~ "Civil Rights Activities",
    activity3 < 470 ~ "Litigation and Legal Aid Activities", 
    activity3 < 510 ~ "Legislative and Political Activities",
    activity3 < 560 ~ "Advocacy",
    activity3 < 600 ~ "Other Activities Directed to Individuals",
    activity3 < 605 ~ "Activities Purposes and Activities",
    activity3 < 450 ~ "Other Purposes and Activities",
    TRUE ~ "NONE" ))




### NTEE Codes 
ntee <- read_csv("https://raw.githubusercontent.com/Nonprofit-Open-Data-Collective/mission-taxonomies/main/NTEE-disaggregated/ntee-crosswalk.csv")

```

## Variable Codings 

### Subsector Code 501c()

| SUBSECCD | Subsector Code 501c()                                                                 |
|--------------|----------------------------------------------------------|
| 01       | 01- Corporations originated under Act of Congress, including Federal Credit Unions    |
| 02       | 02- Title holding corporation for a tax-exempt organization.                          |
| 03       | 03- Religious, educational, charitable, scientific, and literary organizations...     |
| 04       | 04- Civic leagues, social welfare organizations, and local associations of employees  |
| 05       | 05- Labor, agricultural, horticultural organizations. These are eduactional or        |
| 06       | 06- Business leagues, chambers of commerce, real estate boards, etc. formed to impro  |
| 07       | 07- Social and recreational clubs which provide pleasure, recreation, and social      |
| 08       | 08- Fraternal beneficiary societies and associations, with lodges providing for payme |
| 09       | 09- Voluntary employees' beneficiary ass'ns (including fed. employees' voluntary      |
| 10       | 10- Domestic fraternal societies and assoc's-lodges devoting their net earnings to    |
| 11       | 11- Teachers retirement fund associations.                                            |
| 12       | 12- Benevolent life insurance associations, mutual ditch or irrigation companies,     |
| 13       | 13- Cemetery companies, providing burial and incidental activities for members.       |
| 14       | 14- State-chartered credit unions, mutual reserve funds, offering loans to members... |
| 15       | 15- Mutual insurance cos. ar associations, providing insurance to members subst       |
| 16       | 16- Cooperative organizations to finance crop operations, in conjunction with         |
| 17       | 17- Supplemental unemployment benefit trusts, providing payments of suppl. unemp      |
| 18       | 18- Employee funded pension trusts, providing benefits under a pension plan funded by |
| 19       | 19- Post or organization of war veterans.                                             |
| 20       | 20- Trusts for prepaid group legal services, as part of a qual. group legal service   |
| 21       | 21- Black lung trusts, satisfying claims for compensation under Black Lung Acts.      |
| 22       | 22- Multiemployer Pension Plan                                                        |
| 23       | 23- Veterans association formed prior to 1880                                         |
| 24       | 24-Trust described in Section 4049 of ERISA                                           |
| 25       | 25- Title Holding Company for Pensions, etc                                           |
| 26       | 26- State-Sponsored High Risk Health Insurance Organizations                          |
| 27       | 27- State-Sponsored Workers Compensation Reinsurance                                  |
| 40       | 40- Apostolic and religious orgs. - 501(d)                                            |
| 50       | 50- Cooperative Hospital Service Organization - 501(e)                                |
| 60       | 60- Cooperative Service Org. of Operating Educ. Org.- 501(f)                          |
| 70       | 70- Child Care Organization - 501(k)                                                  |
| 71       | 71- Charitable Risk Pool                                                              |
| 80       | 80- Farmers' Cooperatives                                                             |
| 81       | 81- Qualified State-Sponsored Tuition Program                                         |
| 82       | 82- 527 Political Organizations                                                       |
| 90       | 90- 4947(a)(2) Split Interest Trust                                                   |
| 91       | 91- 4947(a)(1) Public Charity (Files 990/990-EZ)                                      |
| 92       | 92- 4947(a)(1) Private Foundations                                                    |
| 93       | 93- 1381(a)(2) Taxable Farmers Cooperative                                            |
| CO       | CO- Unspecified 501(c) Organization Other Than 501(c)(3)                              |






## Initial Explorations 

```{r}
bmf.nccs %>%
  select(LEVEL1, LEVEL2) %>%
  group_by(LEVEL1, LEVEL2) %>%
  summarise(count = n ()) %>%
  ggplot(aes(x= LEVEL1, y = LEVEL2)) +
  geom_tile(aes(fill = log(count, 10))) +
  geom_text(aes(label=count), color = "red")  +
  ggtitle("Level 1 vs Level 2")

  
ggplot(bmf.nccs, aes(x = fct_infreq(LEVEL3))) +
  geom_bar() +
  geom_text(stat='count', aes(label=..count..), 
            vjust=-1, 
            color = "red",
            angle = 30, 
            nudge_x = 0.75)+
  ggtitle("Org Type")

# 501c() type
ggplot(bmf.nccs, aes(x = fct_infreq(as.factor(SUBSECCD)))) +
  geom_bar() +
  geom_text(stat='count', aes(label=..count..), 
            vjust=-1, 
            color = "red",
            angle = 30, 
            nudge_x = 0.75,
            size = 2)+
  ggtitle("501c() Type")

## Filter out 501c(3) to look at the other ones 
bmf.nccs %>%
  filter(SUBSECCD != 3 ) %>%
  ggplot(aes(x = fct_infreq(as.factor(SUBSECCD)))) +
  geom_bar() +
  geom_text(stat='count', aes(label=..count..), 
            vjust=-1, 
            color = "red",
            angle = 30, 
            nudge_x = 0.75,
            size = 2)+
  ggtitle("501c() Type")

#reason for 501c(3) status
bmf.nccs %>%
  select(SUBSECCD, FNDNCD) %>%
  filter(SUBSECCD == 3 ) %>% 
  ggplot(aes(x = fct_infreq(as.factor(FNDNCD)))) +
  geom_bar() +
  geom_text(stat='count', aes(label=..count..), 
            vjust=-1, 
            color = "red",
            angle = 30, 
            nudge_x = 0.75,
            size = 2)+
  ggtitle("Reason for 501c() Status")

#501c(3) status vs LEVEL2
bmf.nccs %>%
  select(SUBSECCD, FNDNCD, LEVEL2) %>%
  filter(SUBSECCD == 3 ) %>%
  group_by(LEVEL2, FNDNCD) %>%
  summarise(count = n ()) %>%
  ggplot(aes(x= LEVEL2, y = FNDNCD)) +
  geom_tile(aes(fill = log(count, 10))) +
  geom_text(aes(label=count), color = "red")  +
  ggtitle("Level 2 vs Reason for 501c(3) Status")

# Reason for 501c(3) Status vs Level 3
bmf.nccs %>%
  select(SUBSECCD, FNDNCD, LEVEL3) %>%
  filter(SUBSECCD == 3 ) %>%
  group_by(LEVEL3, FNDNCD) %>%
  summarise(count = n ()) %>%
  ggplot(aes(x= LEVEL3, y = FNDNCD)) +
  geom_tile(aes(fill = log(count, 10))) +
  geom_text(aes(label=count), color = "red", size=2)  +
  ggtitle("Level 3 vs Reason for 501c(3) Status")

```


## LEVEL3 vs NTEE 

```{r}


dat.NTEE <- 
  bmf.nccs %>%
  #Keep EIN To join later if needed
  select(EIN, NAME, NTEEFINAL, NTEECC, NTEE1, LEVEL4, LEVEL3, LEVEL2) %>%
  left_join(ntee, join_by(NTEEFINAL == NTEE))
  
dat.NTEE%>%
  group_by(LEVEL3, level1) %>%
  summarise(count = n()) %>%
  ggplot(aes(x= LEVEL3, y = level1)) +
  geom_tile(aes(fill = log(count, 10))) +
  geom_text(aes(label=count), color = "red", size = 2)  +
  ggtitle("Level 3 vs Broad Category")

  ## Table of matching codes 
tab.L3.BC <- data.frame(
  L3 = c("AR", "ED", "EN", "HE", "HS", "IN", "MR", "MO", "PB", "RE", "UN"),
  BC = c("ART", "EDU", "ENV", "HEL", "HMS", "IFA", "MMB", "MMB", "PSB", "REL", "UNU")
)

# Table of Codes that Do and do not not Match 
tab.match<- 
  dat.NTEE%>%
  group_by(LEVEL3, level1) %>%
  summarise(count = n ()) %>%
  pivot_wider(names_from = level1, values_from = count) %>%
  relocate(LEVEL3, ART, EDU, ENV, HEL, HMS, IFA, PSB, REL, MMB, UNU, `NA`)  %>%
  replace(is.na(.), 0)

#light blue means the LEVEL 3 and Broad Categories match 
tab.match %>%
  datatable(rownames = FALSE)%>%
  formatStyle(columns = "ART", 
              background = styleEqual(
                sum(dat.NTEE$LEVEL3 == "AR" & dat.NTEE$level1 == "ART", na.rm = TRUE), 
                c("lightblue"))) %>%
  formatStyle(columns = "EDU", 
              background = styleEqual(
                sum(dat.NTEE$LEVEL3 == "ED" & dat.NTEE$level1 == "EDU", na.rm = TRUE), 
                c("lightblue"))) %>%
  formatStyle(columns = "ENV", 
              background = styleEqual(
                sum(dat.NTEE$LEVEL3 == "EN" & dat.NTEE$level1 == "ENV", na.rm = TRUE), 
                c("lightblue"))) %>%
  formatStyle(columns = "HEL", 
              background = styleEqual(
                sum(dat.NTEE$LEVEL3 == "HE" & dat.NTEE$level1 == "HEL", na.rm = TRUE), 
                c("lightblue"))) %>%
  formatStyle(columns = "HMS", 
              background = styleEqual(
                sum(dat.NTEE$LEVEL3 == "HS" & dat.NTEE$level1 == "HMS", na.rm = TRUE), 
                c("lightblue"))) %>%
  formatStyle(columns = "IFA", 
              background = styleEqual(
                sum(dat.NTEE$LEVEL3 == "IN" & dat.NTEE$level1 == "IFA", na.rm = TRUE), 
                c("lightblue"))) %>%
  formatStyle(columns = "MMB", 
              background = styleEqual(
                sum(dat.NTEE$LEVEL3 == "MR" & dat.NTEE$level1 == "MMB", na.rm = TRUE), 
                c("lightblue"))) %>%
  formatStyle(columns = "MMB", 
              background = styleEqual(
                sum(dat.NTEE$LEVEL3 == "MO" & dat.NTEE$level1 == "MMB", na.rm = TRUE), 
                c("lightblue"))) %>%
  formatStyle(columns = "PSB", 
              background = styleEqual(
                sum(dat.NTEE$LEVEL3 == "PB" & dat.NTEE$level1 == "PSB", na.rm = TRUE), 
                c("lightblue"))) %>%
  formatStyle(columns = "REL", 
              background = styleEqual(
                sum(dat.NTEE$LEVEL3 == "RE" & dat.NTEE$level1 == "REL", na.rm = TRUE), 
                c("lightblue"))) %>%
  formatStyle(columns = "UNU", 
              background = styleEqual(
                sum(dat.NTEE$LEVEL3 == "UN" & dat.NTEE$level1 == "UNU", na.rm = TRUE), 
                c("lightblue"))) 
  


```


Comments:

-   "AR", "ED", "EN", "HE", "IN", "MO", "MR", "PB", "RE", "UN" and "-"/NA, have clear NTEE Equivalents.

-   The ZA, ..., ZF need to be investigated.

    -   ZA, ZB are fairly evenly spread across broad categories

    -   ZC seems to be concentrated in PSB, but has a lot in other broad categories as well.

    -   ZD, ZE, ZF seem to be concentrated in PSB, but


#### Level 3 == "ZA" or "ZB"

```{r}

#LEVEL3 vs NTEE org type (level 5)
dat.NTEE%>%
  filter(str_sub(LEVEL3, 1, 1) == "Z") %>%
  group_by(LEVEL3, level5) %>%
  summarise(count = n ()) %>%
  pivot_wider(names_from = level5, values_from = count) %>%
  datatable()

```

Comments: 

- LEVEL3 Code "ZA" is NTEE Common Code MS (11- Monetary Support - Single Org) 

- LEVEL3 Code "ZB" is NTEE Common Code MM (12- Monetary Support - Multiple Orgs)

Let's explore what's going on with ZC - ZF

#### LEVEL3 == "ZA"

```{r}
### LEVEL3 == "ZA" vs Major Group
dat.NTEE%>%
  filter(LEVEL3 == "ZA") %>%
  group_by(LEVEL3, level2) %>%
  summarise(count = n ())  %>% 
  arrange(desc(count)) %>%
  datatable()
```

Comments: 

- Most common is B, but there are thousands outside of group B. 

- I don't think ZA corresponds to a specific major group. I think it only corresponds to specialty org type Monetary Support - Single Org

#### LEVEL3 == "ZB"

```{r}
### LEVEL3 == "ZB" vs Major Group
dat.NTEE%>%
  filter(LEVEL3 == "ZB") %>%
  group_by(LEVEL3, level2) %>%
  summarise(count = n ())  %>% 
  arrange(desc(count)) %>%
  datatable()
```

Comments: 

- Most common is B, but there are thousands outside of group B. 

- I don't think ZA corresponds to a specific major group. I think it only corresponds to specialty org type Monetary Support - Multiple Orgs


#### LEVEL3 == "ZC"

```{r}
### LEVEL3 == "ZC" vs Major Group
dat.NTEE%>%
  filter(LEVEL3 == "ZC") %>%
  group_by(LEVEL3, level2) %>%
  summarise(count = n ())  %>% 
  arrange(desc(count)) %>%
  datatable()

dat.NTEE%>%
  filter(LEVEL3 == "ZC") %>%
  group_by(LEVEL3,level4, level4.label) %>%
  summarise(count = n ())  %>% 
  arrange(desc(count)) %>%
  datatable()
```

Comments:

-   Mostly Major Group T: Philanthropy, Voluntarism & Grantmaking Foundations 

- This makes sense since "ZC" is Private grantmaking foundations


### LEVEL3 == "ZD"

```{r}
### LEVEL3 == "ZC" vs Major Group
dat.NTEE%>%
  filter(LEVEL3 == "ZD") %>%
  group_by(LEVEL3, level3) %>%
  summarise(count = n ()) 

dat.NTEE%>%
  filter(LEVEL3 == "ZD") %>%
  group_by(LEVEL3,level4, level4.label) %>%
  summarise(count = n ())  %>% 
  arrange(desc(count)) %>%
  datatable()
```


Comments : 

- All of these are in T (Philanthropy, Voluntarism & Grantmaking Foundations)

- All of these are T30 (Public Foundations) and T31 (Community Foundations).


### LEVEL3 == "ZE"

```{r}
### How many ZE orgs
dat.NTEE%>%
  filter(LEVEL3 == "ZE") %>%
  group_by(LEVEL3) %>%
  summarise(count = n ()) 

### ZE vs major group
dat.NTEE%>%
  filter(LEVEL3 == "ZE") %>%
  group_by(LEVEL3, level4, level4.label) %>%
  summarise(count = n ())  %>% 
  arrange(desc(count)) %>%
  datatable()
```

Comments:

-   Vast majority of these are T70 (Philanthropy, Voluntarism & Grantmaking Foundations) - (Organizations that administer centralized fundraising efforts and then in turn distribute the contributed funds to several nonprofit agencies.)



### Level3 == "ZF"

```{r}
### How many ZF orgs
dat.NTEE%>%
  filter(LEVEL3 == "ZE") %>%
  group_by(LEVEL3) %>%
  summarise(count = n ()) 

### ZF vs major group
dat.NTEE%>%
  filter(LEVEL3 == "ZF") %>%
  group_by(LEVEL3, level4, level4.label) %>%
  summarise(count = n ())  %>% 
  arrange(desc(count))%>%
  datatable
```



Comments:

- all unclassified trusts



### Z\* conclusions

Updated table of most likely LEVEL3 to NTEE equvilants.

| LEVEL3 Code | LEVEL 3 name                        | Most likely NTEE Code Equivalent                                                           |  Comments                                                                 |
|--------------|--------------|------------------|--------------|--------------|
| ZA          | Single organization support         | MS - 11 (Monetary Support - Single Organization)                                                                  | This LEVEL3 code doesn't seem to be particular to a specific major group |
| ZB          | Fundraising within NTEE major group | MM - Potentially 12 (Monetary Support - Multiple Organizations)                                                              | This LEVEL3 code doesn't seem to be particular to a specific major group |
| ZC          | Private grantmaking foundations     | T2\* (Private Grantmaking Foundations). There are some outside of this NTEE Code though    |            |                           
| ZD          | Public foundations                  | T3\* (Private Operating Foundations).                                                                                                        |. |
| ZE          | General fundraising                 | T70 (Federated Giving Programs)                                                             |                                                                          |
| ZF          | Other Supporting Public Benefit     | T90(Named Trusts N.E.C.), T99 (Philanthropy, Voluntarism & Grantmaking Foundations N.E.C.)                                    |                                                                          |









## Status 

* 01 Unconditional Exemption

* 02 Conditional Exemption

* 12 Trust described in section 4947(a)(2) of the IR Code

* 25 Organization terminating its private foundation status under section 507(b)(1)(B) of the Code

```{r}
table(bmf.irs$STATUS)
```

### Let's first focus on STATUS == 12 

```{r}
status12 <- bmf.irs %>% 
  filter(STATUS == 12) %>% 
  left_join(ntee, join_by(NTEE_CD == NTEE ))


table(status12$level1, useNA = "always")


table(status12$FOUNDATION)

table(status12$SUBSECTION)

table(status12$CLASSIFICATION, status12$SUBSECTION)

table(status12$PF_FILING_REQ_CD) #1 is required to complete 990PF


#those that are not required to file a PF, what are they required to file?
status12 %>% 
  filter(PF_FILING_REQ_CD == 0) %>% 
  pull(FILING_REQ_CD) %>% 
  table()



```

Comments: 

- Most NTEE codes in STATUS == 12 are NA

- Most are not 501c(3)'s, some are 501c3 - Private non-operating foundation

- Most are in classification 1, primarily subsection 91 (Non-Exempt Charitable Trusts	?) and 92 (4947(a)(1) - Private Foundation (Form 990PF Filer))

- Most have PF FILING REQUIREMENT CODE == 1 (990-PF return), but some are not required to. 

- It seems to be that most status == 12 (trust) are private foundations, which makes sense. 

  - But there are ~800 who aren't required to file 990PF. Of those, they are either requried to file 990/990EZ/990-N or not required to file at all. 


- The organization code doesn't make sense 

```{r}
table(status12$ORGANIZATION)
```


status == 12 says they are a trust, but their organization code says only 539 of them are trusts (organization == 2) 



### Next, Status == 02

```{r}
status2 <- bmf.irs %>% 
  filter(STATUS == 2) %>% 
  left_join(ntee, join_by(NTEE_CD == NTEE ))


table(status2$NTEE_CD, useNA = "always")

table(status2$FOUNDATION)

table(status2$SUBSECTION)

table(status2$SUBSECTION, status2$CLASSIFICATION)


# table(status2$SUBSECTION, status2$class1)
# table(status2$SUBSECTION, status2$class2)
# table(status2$SUBSECTION, status2$class3)
# table(status2$SUBSECTION, status2$class4)
# table(status2$SUBSECTION, status2$class7)


table(status2$FILING_REQ_CD)

table(status2$activity1.group)
table(status2$activity2.group)
table(status2$activity3.group)


```


Comments: 

- Foundation codes are either 0 or 10 (church)

- Most common subsection/classification is 501c(3) - classification 7 (church)

- Filing requirement code 6 (not required - church) is most common, although there are a lot in F_REQ_Code 0 (not required), 1 (990 or 990EZ), 2(990-N)

- most activity code groups are religious activities

- Seems to be most subsection 02 are churches/religious orgs




