---
title: "Governance Workflow"
author: "Olivia Beck"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(data.table)

source("governance/pkg-funcs/get-features.R")
source("governance/pkg-funcs/get-scores.R")

```

This is a workflow of the governance index functions that will be used in the governance package. 

## 1. Get Input data 

We assume the user already has the relevant data downloaded from the 990's. Here is us getting that data: 

```{r}
#set the years you are interesteed in
years = 2018

```


### Part IV data 

```{r}
### Part IV 

#initialize data
dat_4 <-  vector(mode = "list", length = length(years))
#get columns I want 
keep_cols_part4 <- c("OBJECTID", "URL", "RETURN_VERSION", "ORG_EIN", 
                     "F9_04_AFS_IND_X", "F9_04_AFS_CONSOL_X",
                     "F9_04_BIZ_TRANSAC_DTK_X", "F9_04_BIZ_TRANSAC_DTK_FAM_X", "F9_04_BIZ_TRANSAC_DTK_ENTITY_X",
                     "F9_04_CONTR_NONCSH_MT_25K_X", 
                     "F9_04_CONTR_ART_HIST_X")

#read in the data
for(i in 1:length(years)){
  link <-  paste0("https://nccs-efile.s3.us-east-1.amazonaws.com/parsed/F9-P04-T00-REQUIRED-SCHEDULES-", years[i], ".csv")
  temp <- fread(link, select = keep_cols_part4) 
  dat_4[[i]] <- temp
}

#clean up data
dat_all_4 <- 
  rbindlist(dat_4) %>% 
  mutate( year = as.numeric(substr(RETURN_VERSION, 1, 4)))%>% 
  filter(year <= max(years))
```

### Part VI Data

```{r}
### Part VI 

#initialize data
dat_6 <-  vector(mode = "list", length = length(years))
#get columns I want 
keep_cols_part6 <- c("OBJECTID", "URL", "RETURN_VERSION", "ORG_EIN", 
                     "F9_06_INFO_SCHED_O_X", 
                     "F9_06_GVRN_NUM_VOTING_MEMB",
                     "F9_06_GVRN_NUM_VOTING_MEMB_IND", 
                     "F9_06_GVRN_DTK_FAMBIZ_RELATION_X", 
                     "F9_06_GVRN_DELEGATE_MGMT_DUTY_X",
                     "F9_06_GVRN_CHANGE_DOC_X", 
                     "F9_06_GVRN_ASSET_DIVERSION_X",
                     "F9_06_GVRN_MEMB_STCKHLDR_X", 
                     "F9_06_GVRN_PERS_ELECT_MEMB_X",
                     "F9_06_GVRN_DECISION_APPROV_X",
                     "F9_06_GVRN_DOC_GVRN_BODY_X",
                     "F9_06_GVRN_DOC_COMMITTEE_X",
                     "F9_06_POLICY_CHPTR_AFFIL_X",
                     "F9_06_POLICY_CHPTR_REFERENCE_X",
                     "F9_06_POLICY_FORM990_GVRN_BODY_X",
                     "F9_06_POLICY_COI_X",
                     "F9_06_POLICY_COI_DISCLOSURE_X",
                     "F9_06_POLICY_COI_MONITOR_X",
                     "F9_06_POLICY_WHSTLBLWR_X",
                     "F9_06_POLICY_DOC_RETENTION_X",
                     "F9_06_POLICY_COMP_PROCESS_CEO_X",
                     "F9_06_POLICY_COMP_PROCESS_OTH_X",
                     "F9_06_POLICY_JV_X",
                     "F9_06_POLICY_JV_PROC_X",
                     "F9_06_DISCLOSURE_AVBL_OTH_X",
                     "F9_06_DISCLOSURE_AVBL_OTH_WEB_X",
                     "F9_06_DISCLOSURE_AVBL_REQUEST_X",
                     "F9_06_DISCLOSURE_AVBL_OWN_WEB_X"
                     )

#read in data
for(i in 1:length(years)){
  link <-  paste0("https://nccs-efile.s3.us-east-1.amazonaws.com/parsed/F9-P06-T00-GOVERNANCE-", years[i], ".csv")
  temp <- fread(link, select = keep_cols_part6) 
  dat_6[[i]] <- temp
}

#clean up data
dat_all_6 <- 
  rbindlist(dat_6) %>% 
  mutate( year = as.numeric(substr(RETURN_VERSION, 1, 4))) %>% 
  filter(year <= max(years))
```


### Part XII Data

```{r}
### Part XII

#initialize data
dat_12 <-  vector(mode = "list", length = length(years))
#Keep all columns for part XII


for(i in 1:length(years)){
  link <-  paste0("https://nccs-efile.s3.us-east-1.amazonaws.com/parsed/F9-P12-T00-FINANCIAL-REPORTING-", years[i], ".csv")
  temp <- fread(link) 
  dat_12[[i]] <- temp
}

#clean up data
dat_all_12 <- 
  rbindlist(dat_12) %>% 
  mutate( year = as.numeric(substr(RETURN_VERSION, 1, 4)))%>% 
  filter(year <= max(years))


```

### Schedule M Data

```{r}
#initialize data
dat_M <-  vector(mode = "list", length = length(years))
#get columns I want 
keep_cols_partM <- c("OBJECTID", "URL", "RETURN_VERSION", "ORG_EIN", 
                     "SM_01_REVIEW_PROCESS_UNUSUAL_X")

for(i in 1:length(years)){
  link <-  paste0("https://nccs-efile.s3.us-east-1.amazonaws.com/parsed/SM-P01-T00-NONCASH-CONTRIBUTIONS-", years[i], ".csv")
  temp <- fread(link, select = keep_cols_partM) 
  dat_M[[i]] <- temp
}


#clean up data
dat_all_M <- 
  rbindlist(dat_M) %>% 
  mutate( year = as.numeric(substr(RETURN_VERSION, 1, 4)))%>% 
  filter(year <= max(years))
```


### Merge all parts 

```{r}
dat_govern <-
  dat_all_4 %>% 
  merge(dat_all_6, by = c("OBJECTID", "URL", "RETURN_VERSION", "ORG_EIN", "year")) %>% 
  merge(dat_all_12, by = c("OBJECTID", "URL", "RETURN_VERSION", "ORG_EIN", "year")) %>% 
  merge(dat_all_M, by = c("OBJECTID", "URL", "RETURN_VERSION", "ORG_EIN", "year")) %>% 
  filter(RETURN_TYPE != "990EZ")
```

### Only take a sample for demonstration purposes

```{r}
set.seed(978345)
dat_testing <- 
  dat_govern  %>% 
  sample_n(300)
```



## Step 2: Get Features Matrix 

Use the `get_features` function to clean the data and transform it into a features matrix. 

```{r}
features_testing <- get_features(dat_testing)
head(features_testing)
```


## Step 3: Calculate the Scores 

Use the `get_scores` function to get the governance scores. 

Issue: all the socres are NaN. I do not know why this is happening. 

```{r}
scores_testing <- get_scores(features_testing)
head(scores_testing)
hist(scores_testing$total.score)
```

## When issues can occur

If 1 column has entirely 1's or 0's, then the `get_scores` function will return NaN. This is because there is a matrix inversion in the solution process, which cannot be done with non full rank matrices. 

### Generating Random Data for features matrix - bad example

```{r}
#Random Data is fine 
new.dat <- matrix(sample(0:1, 200*12, replace = T), ncol = 12 , nrow = 200)
new.dat <- as.data.frame(new.dat)
colnames(new.dat) <- colnames(features_testing[, -1])

new.dat$P4_LINE_29_30 <- 0

new.scores <- get_scores(new.dat)
head(new.scores)

```
