---
title: "Script to Make Board Governance Data"
author: "Olivia Beck"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(data.table)

```

## Parts of 990 to Read In

### Part IV

12a. Whether obtained separate, independent audited financial statements
(y/n)

12b. Whether organization included in consolidated, independent audited
financial statements (y/n)

28a. Whether the organization a party to a business transaction with a
current or former officer, director, trustee, or key employee (y/n)

28b. Whether the organization a party to a business transaction to with
a family member of a current or former officer, director, trustee, or
key employee (y/n)

28c. Whether the organization a party to a business transaction to with
an entity of which a current or former officer, director, trustee, or
key employee was an officer, director, trustee, or direct or indirect
owner (y/n)

29. Did the organization receive more than 25,000 in non-cash
    contributions? If yes, complete schedule M. (y/n)

30. Did the organization receive contributions of art, historical
    treasures, or other similar assets or qualified conservation
    contributions? If yes, complete schedule M. (y/n)

\*Note that if any "yes" to 29 or 30, org is required to indicate
whether it has a "gift acceptance policy" in Schedule M, item 31 (y/n)

#### Part IV Code Book

+----------------------+----------------------+----------------------+
| Line                 | Variable             | Levels               |
+======================+======================+======================+
| Line 12A, 12B, 28A,  | P4_LINE_12A, ...,    | -   "yes" - the      |
| 28B, 28C, 29, 30     | P4_LINE_30           |     answer to this   |
|                      |                      |     line is yes      |
|                      |                      |                      |
|                      |                      | -   "no" - the       |
|                      |                      |     answer to this   |
|                      |                      |     line is no       |
+----------------------+----------------------+----------------------+

### Part VI

1a. Number of Voting Members

1b. Number of independent voting members

1.schedule-O. If there are material differences in voting rights among
members of the governing body, or if the governing body delegated broad
authority to an executive committee or similar committee, explain on
Schedule O.

2.  Family relationship or business relationship (y/n)
3.  Whether used a management company or other person to perform any
    management duties (y/n)
4.  Whether made any significant changes to governing documents (y/n)
    5.Whether any significant diversion of assets (y/n)
5.  Whether have members or stockholders (y/n)

7a. Whether have any persons who had power to elect or appoint members
(y/n) 7b. Whether decisions reserved to members other than governing
body (y/n)

8a.Whether documented the meetings by the governing body (y/n) 8b.
Whether documented the meetings by each committee (y/n)

10a. Whether have local chapters, branches, or affiliates (y/n) 10b. If
yes, whether have a written policies governing the activities for local
chapters, branches, or affiliates (y/n)

11a. Whether provided a copy of Form 990 to all members before filing
the form (y/n)

12a. Whether have a written conflict of interest policy (y/n) 12b.
Whether officers, directors, key employees required to disclose annually
interests that could give rise to conflicts? (y/n) 12c. Whether
regularly and consistently monitor and enforce compliance with the
policy (y/n)

13. Whether have a written whistleblower policy\* (y/n)
14. Whether have a written document retention and destruction policy\*
    (y/n)

15a. Whether have a process for determining compensation of the CEO,
executive director, or top management officials include a review and
approval by independent persons, comparability data, and deliberation
decision (y/n) 15b. Whether have a process for determining compensation
of other officers or key employees include a review and approval by
independent persons, comparability data, and deliberation decision (y/n)

16a. Whether invest in, contribute assets to, or participate in a joint
venture or similar arrangement (y/n) 16b. If yes, whether have a written
policy or procedure to evaluate its participation in joint venture or
similar arrangements (y/n)

18. Whether make Forms 1023 (or 1024), 990, 990-T available for public
    inspection in any of the following forms: Own website, another's
    website, upon request, other

#### Part VI Code Book

+----------------------+----------------------+----------------------+
| Line Number          | Variable Name        | Levels               |
+======================+======================+======================+
| Line 1A, 1B          | P6_LINE_1A,          | integer value        |
|                      | P6_LINE_1B           |                      |
+----------------------+----------------------+----------------------+
| Line 2, 3, 4, 5, 6,  | P6_LINE_2, .... ,    | -   "yes" - the      |
| 7A, 7B, 8A, 8B, 10A, | P6_LINE_16B          |     answer to this   |
| 10B, 11A, 12A, 12B,  |                      |     line is yes      |
| 12C, 13, 14, 15A,    |                      |                      |
| 15B, 16A, 16B        |                      | -   "no" - the       |
|                      |                      |     answer to this   |
|                      |                      |     line is no       |
+----------------------+----------------------+----------------------+
| Line 18              | P6_LINE_18           | -   "own_website" -  |
|                      |                      |     the own website  |
|                      |                      |     box on line 18   |
|                      |                      |     is checked       |
|                      |                      |                      |
|                      |                      | -                    |
|                      |                      |   "others_website" - |
|                      |                      |     the another's    |
|                      |                      |     website box on   |
|                      |                      |     line 18 is       |
|                      |                      |     checked          |
|                      |                      |                      |
|                      |                      | -   "request" - the  |
|                      |                      |     upon request     |
|                      |                      |     website box on   |
|                      |                      |     line 18 is       |
|                      |                      |     checked          |
|                      |                      |                      |
|                      |                      | -   "other" - the    |
|                      |                      |     other box on     |
|                      |                      |     line 18 is       |
|                      |                      |     checked          |
+----------------------+----------------------+----------------------+
| Line 1 Schedule O    | P6_SCHEDULE_O        | -   "X" - there is a |
|                      |                      |     Schedule O       |
|                      |                      |     attached to this |
|                      |                      |     return with      |
|                      |                      |     notes from Part  |
|                      |                      |     VI Line 1        |
|                      |                      |                      |
|                      |                      | -   NA - there is no |
|                      |                      |     Schedule O       |
|                      |                      |     attached to this |
|                      |                      |     return with      |
|                      |                      |     notes from Part  |
|                      |                      |     VI Line 1        |
+----------------------+----------------------+----------------------+

### Part XII

1.  Accounting method used to prepare the Form 990 in any of the
    following forms: cash, accrual, other

2a. Were the financial statements compiled or reviewed by an independent
accountant\* (y/n)

2b. Were the financial statements audited by an independent accountant\*
(y/n)

2c. If yes to 2a or 2b, whether have a committee that assumes
responsibility for oversight of the audit, review, or compilation of its
financial statements and selection of an independent accountant? (y/n)

3a. Was required to undergo an audit or audits (y/n)

3b. If yes, did undergo the required audit or audits (y/n)

#### Part 12 Code Book

+---------------------+---------------------+---------------------+
| Line Number         | Variable Name       | Levels              |
+=====================+=====================+=====================+
| Line 1              | P12_LINE_01         | -   "cash" - cash   |
|                     |                     |     accounting      |
|                     |                     |     method          |
|                     |                     |                     |
|                     |                     | -   "accrual" -     |
|                     |                     |     accrual         |
|                     |                     |     accounting      |
|                     |                     |     method          |
|                     |                     |                     |
|                     |                     | -   "other" - other |
|                     |                     |     accounting      |
|                     |                     |     method          |
+---------------------+---------------------+---------------------+
| Line 1              | P12_L               | If P12_LINE_01 is   |
|                     | I                   | "other", then this  |
|                     | NE_01_OTHER_EXPLAIN | is what is written  |
|                     |                     | for other           |
+---------------------+---------------------+---------------------+
| Line 2a             | P12_LINE_2A         | -   "no" - answer   |
|                     |                     |     to question 2a  |
|                     |                     |     is no           |
|                     |                     |                     |
|                     |                     | -   "separate" -    |
|                     |                     |     answer to       |
|                     |                     |     question 2a is  |
|                     |                     |     yes and the box |
|                     |                     |     for "separate   |
|                     |                     |     basis" is       |
|                     |                     |     checked         |
|                     |                     |                     |
|                     |                     | -   "colsol" -      |
|                     |                     |     answer to       |
|                     |                     |     question 2a is  |
|                     |                     |     yes and the box |
|                     |                     |     for             |
|                     |                     |     "consolidated   |
|                     |                     |     basis" is       |
|                     |                     |     checked         |
|                     |                     |                     |
|                     |                     | -   "both" - answer |
|                     |                     |     to question 2a  |
|                     |                     |     is yes and the  |
|                     |                     |     box for "both   |
|                     |                     |     consolidated    |
|                     |                     |     and separate    |
|                     |                     |     basis" is       |
|                     |                     |     checked         |
+---------------------+---------------------+---------------------+
| Line 2b             | P12_LINE_2B         | -   "no" - answer   |
|                     |                     |     to question 2a  |
| Notes: Something is |                     |     is no           |
| wrong about this    |                     |                     |
| coding. There are   |                     | -   "separate" -    |
| many F              |                     |     answer to       |
| 9                   |                     |     question 2b is  |
| _12_FINSTAT_AUDIT_X |                     |     yes and the box |
| == FALSE, But       |                     |     for "separate   |
| F9_12_FIN           |                     |     basis" is       |
| S                   |                     |     checked         |
| TAT_AUDIT_CONSOL_X, |                     |                     |
| F9_12\_             |                     | -   "colsol" -      |
| F                   |                     |     answer to       |
| INSTAT_AUDIT_SEP_X, |                     |     question 2b is  |
| or F9_12\_          |                     |     yes and the box |
| F                   |                     |     for             |
| INSTAT_AUDIT_BOTH_X |                     |     "consolidated   |
| is checked          |                     |     basis" is       |
|                     |                     |     checked         |
|                     |                     |                     |
|                     |                     | -   "both" - answer |
|                     |                     |     to question 2b  |
|                     |                     |     is yes and the  |
|                     |                     |     box for "both   |
|                     |                     |     consolidated    |
|                     |                     |     and separate    |
|                     |                     |     basis" is       |
|                     |                     |     checked         |
+---------------------+---------------------+---------------------+
| Line 2C             | P12_LINE_2C         | -   "no" - answer   |
|                     |                     |     to question 2c  |
|                     |                     |     is no           |
|                     |                     |                     |
|                     |                     | -   "yes" - answer  |
|                     |                     |     to question 2c  |
|                     |                     |     is no           |
+---------------------+---------------------+---------------------+
| Line 3A             | P12_LINE_3A         | -   "no" - answer   |
|                     |                     |     to question 3a  |
|                     |                     |     is no           |
|                     |                     |                     |
|                     |                     | -   "yes" - answer  |
|                     |                     |     to question 3a  |
|                     |                     |     is no           |
+---------------------+---------------------+---------------------+
| Line 3B             | P12_LINE_3B         | -   "no" - answer   |
|                     |                     |     to question 3b  |
|                     |                     |     is no           |
|                     |                     |                     |
|                     |                     | -   "yes" - answer  |
|                     |                     |     to question 3b  |
|                     |                     |     is no           |
+---------------------+---------------------+---------------------+

## Load In Needed Data

```{r}
years = 2009:2013

```

### Part IV

```{r}
### Part IV 

#initialize data
dat_4 <-  vector(mode = "list", length = length(years))
#get columns I want 
keep_cols_part4 <- c("OBJECTID", "URL", "RETURN_VERSION", "ORG_EIN", 
                     "F9_04_AFS_IND_X", "F9_04_AFS_CONSOL_X",
                     "F9_04_BIZ_TRANSAC_DTK_X", "F9_04_BIZ_TRANSAC_DTK_FAM_X", "F9_04_BIZ_TRANSAC_DTK_ENTITY_X",
                     "F9_04_CONTR_NONCSH_MT_25K_X", 
                     "F9_04_CONTR_ART_HIST_X")

for(i in 1:length(years)){
  link <-  paste0("https://nccs-efile.s3.us-east-1.amazonaws.com/parsed/F9-P04-T00-REQUIRED-SCHEDULES-", years[i], ".csv")
  temp <- fread(link, select = keep_cols_part4) 
  colnames(temp) <- c("OBJECTID", "URL", "RETURN_VERSION", "ORG_EIN",
                     "P4-LINE-12A", "P4-LINE-12B",
                     "P4-LINE-28A", "P4-LINE-28B", "P4-LINE-28C",
                     "P4-LINE-29", 
                     "P4-LINE-30")
  dat_4[[i]] <- temp
}

#clean up data
dat_all_4 <- 
  rbindlist(dat_4) %>% 
  mutate( year = as.numeric(substr(RETURN_VERSION, 1, 4)))

dat_all_4 <-
  dat_all_4 %>% 
  filter(year <= max(years))

#standardize yes/no answers
dat_all_4 <- 
  dat_all_4 %>% 
  mutate(across(starts_with("P4"), 
                ~ case_when(
                  . %in% c("true", "1") ~ "yes",
                  . %in% c("false", "0") ~ "no", 
                  TRUE ~ .
                ))) 

```

### Part VI

```{r}
### Part VI 

#initialize data
dat_6 <-  vector(mode = "list", length = length(years))
#get columns I want 
keep_cols_part6 <- c("OBJECTID", "URL", "RETURN_VERSION", "ORG_EIN", 
                     "F9_06_INFO_SCHED_O_X", 
                     "F9_06_GVRN_NUM_VOTING_MEMB",
                     "F9_06_GVRN_NUM_VOTING_MEMB_IND", 
                     "F9_06_GVRN_DTK_FAMBIZ_RELATION_X", 
                     "F9_06_GVRN_DELEGATE_MGMT_DUTY_X",
                     "F9_06_GVRN_CHANGE_DOC_X", 
                     "F9_06_GVRN_ASSET_DIVERSION_X",
                     "F9_06_GVRN_MEMB_STCKHLDR_X", 
                     "F9_06_GVRN_PERS_ELECT_MEMB_X",
                     "F9_06_GVRN_DECISION_APPROV_X",
                     "F9_06_GVRN_DOC_GVRN_BODY_X",
                     "F9_06_GVRN_DOC_COMMITTEE_X",
                     "F9_06_POLICY_CHPTR_AFFIL_X",
                     "F9_06_POLICY_CHPTR_REFERENCE_X",
                     "F9_06_POLICY_FORM990_GVRN_BODY_X",
                     "F9_06_POLICY_COI_X",
                     "F9_06_POLICY_COI_DISCLOSURE_X",
                     "F9_06_POLICY_COI_MONITOR_X",
                     "F9_06_POLICY_WHSTLBLWR_X",
                     "F9_06_POLICY_DOC_RETENTION_X",
                     "F9_06_POLICY_COMP_PROCESS_CEO_X",
                     "F9_06_POLICY_COMP_PROCESS_OTH_X",
                     "F9_06_POLICY_JV_X",
                     "F9_06_POLICY_JV_PROC_X",
                     "F9_06_DISCLOSURE_AVBL_OTH_X",
                     "F9_06_DISCLOSURE_AVBL_OTH_WEB_X",
                     "F9_06_DISCLOSURE_AVBL_REQUEST_X",
                     "F9_06_DISCLOSURE_AVBL_OWN_WEB_X"
                     )

for(i in 1:length(years)){
  link <-  paste0("https://nccs-efile.s3.us-east-1.amazonaws.com/parsed/F9-P06-T00-GOVERNANCE-", years[i], ".csv")
  temp <- fread(link, select = keep_cols_part6) 
  colnames(temp) <- c("OBJECTID", "URL", "RETURN_VERSION", "ORG_EIN",
                     "P6_SCHEDULE_O", 
                     "P6_LINE_1A",
                     "P6_LINE_1B", 
                     "P6_LINE_2", 
                     "P6_LINE_3",
                     "P6_LINE_4", 
                     "P6_LINE_5",
                     "P6_LINE_6",
                     "P6_LINE_7A",
                     "P6_LINE_7B",
                     "P6_LINE_8A",
                     "P6_LINE_8B",
                     "P6_LINE_10A",
                     "P6_LINE_10B",
                     "P6_LINE_11A",
                     "P6_LINE_12A",
                     "P6_LINE_12B",
                     "P6_LINE_12C",
                     "P6_LINE_13",
                     "P6_LINE_14",
                     "P6_LINE_15A",
                     "P6_LINE_15B",
                     "P6_LINE_16A",
                     "P6_LINE_16B",
                     "P6_LINE_18_other", "P6_LINE_18_other_web", "P6_LINE_18_req", "P6_LINE_18_own_web")
  dat_6[[i]] <- temp
}

#clean up data
dat_all_6 <- 
  rbindlist(dat_6) %>% 
  mutate( year = as.numeric(substr(RETURN_VERSION, 1, 4)))

dat_all_6 <-
  dat_all_6 %>% 
  filter(year <= max(years))



dat_all_6 <- 
  dat_all_6 %>% 
  #standardize yes/no answers
  mutate(across(paste0("P6_LINE_", c(2:6, "7A", "7B", "8A", "8B", "10A", "10B", "11A", "12A", "12B", "12C", 13, 14, "15A", "15B", "16A", "16B")) , 
                ~ case_when(
                  . %in% c("true", "1") ~ "yes",
                  . %in% c("false", "0") ~ "no", 
                  TRUE ~ .
                )) ) %>% 
 # Line 18
  mutate(P6_LINE_18 = case_when(
    P6_LINE_18_own_web == "X" ~ "own_website", 
    P6_LINE_18_other_web == "X" ~ "others_website",
    P6_LINE_18_req == "X" ~ "request",
    P6_LINE_18_other == "X" ~ "other"
  )) %>%
  select(-c(P6_LINE_18_own_web,P6_LINE_18_other_web,P6_LINE_18_req, P6_LINE_18_other )) 



```

### Part XII

```{r}
### Part XII

#initialize data
dat_12 <-  vector(mode = "list", length = length(years))
#Keep all columns for part XII


for(i in 1:length(years)){
  link <-  paste0("https://nccs-efile.s3.us-east-1.amazonaws.com/parsed/F9-P12-T00-FINANCIAL-REPORTING-", years[i], ".csv")
  temp <- fread(link) 
  dat_12[[i]] <- temp
}




#clean up data
dat_all_12 <- 
  rbindlist(dat_12) %>% 
  mutate( year = as.numeric(substr(RETURN_VERSION, 1, 4)))

dat_all_12 <-
  dat_all_12 %>% 
  filter(year <= max(years))

## Clean Up Variables to make them more readable
dat_all_12 <-
  dat_all_12 %>% 
  ## Line 1
  mutate(P12_LINE_01 = case_when(
    !is.na(F9_12_FINSTAT_METHOD_ACC_OTH) ~ "other",
    F9_12_FINSTAT_METHOD_ACC_ACCRU_X == "X" ~ "accrual",
    F9_12_FINSTAT_METHOD_ACC_CASH_X == "X" ~ "cash"
  )) %>% 
  mutate(P12_LINE_01_OTHER_EXPLAIN = 
           ifelse(P12_LINE_01 == "other", 
                  F9_12_FINSTAT_METHOD_ACC_OTH,
                  NA)) %>% 
  ### Line 2a
  mutate(temp_2a = case_when(
    F9_12_FINSTAT_COMPILE_X %in% c("true", "1") ~ "1",
    TRUE ~ "0",
  )) %>% 
  mutate(P12_LINE_2A = case_when(
    temp_2a == "0" ~ "no",
    F9_12_FINSTAT_COMPILE_SEP_X == "X" ~ "separate",
    F9_12_FINSTAT_COMPILE_CONSOL_X == "X" ~ "consol",
    F9_12_FINSTAT_COMPILE_BOTH_X == "X" ~ "both"
  )) %>% 
  ### Line 2b
  mutate(temp_2b = case_when(
    F9_12_FINSTAT_AUDIT_X %in% c("true", "1") ~ "1",
    TRUE ~ "0",
  ))  %>% 
  mutate(P12_LINE_2B = case_when(
    temp_2a == "0" ~ "no",
    F9_12_FINSTAT_AUDIT_SEP_X == "X" ~ "separate",
    F9_12_FINSTAT_AUDIT_CONSOL_X == "X" ~ "consol",
    F9_12_FINSTAT_AUDIT_BOTH_X == "X" ~ "both"
  ))%>%
  ### Line 2c
  mutate(P12_LINE_2C = case_when(
    F9_12_FINSTAT_AUDIT_COMMITTEE_X %in% c("true", "1") ~ "yes",
    TRUE ~ "no",
  ))  %>%  
 ### Line 3a
  mutate(P12_LINE_3A = case_when(
    F9_12_FINSTAT_FED_AUDIT_REQ_X %in% c("true", "1") ~ "yes",
    TRUE ~ "no",
  )) %>%
  ### Line 3b
  mutate(P12_LINE_3B = case_when(
    F9_12_FINSTAT_FED_AUDIT_PERF_X %in% c("true", "1") ~ "yes",
    TRUE ~ "no",
  ))  %>% 
  #ony keeping releveant variables
  select(OBJECTID,URL, RETURN_VERSION,ORG_EIN,RETURN_TYPE,TAX_YEAR,
         P12_LINE_01,P12_LINE_01_OTHER_EXPLAIN, 
         P12_LINE_2A, P12_LINE_2B, P12_LINE_2C,
         P12_LINE_3A, P12_LINE_3B) %>%
  ### rename year
  rename(year = TAX_YEAR)


```

## Merge all parts

```{r}
#merge 
#filter out all EZ's

dat_govern <-
  dat_all_4 %>% 
  merge(dat_all_6, by = c("OBJECTID", "URL", "RETURN_VERSION", "ORG_EIN", "year")) %>% 
  merge(dat_all_12, by = c("OBJECTID", "URL", "RETURN_VERSION", "ORG_EIN", "year")) %>% 
  filter(RETURN_TYPE != "990EZ")




save(dat_govern, file = "governance/governance.Rdata")

```
