---
title: "Board Of Governors Explore"
author: "Olivia Beck"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(data.table)
library(viridis)

```

## Load In Part VII Data

```{r}
years <- 2009:2013
dat_7 <-  vector(mode = "list", length = length(years))
keep_cols_7 <- c("EIN", "TAXYR", "F9_07_COMP_DTK_TITLE", "first_name", "suffix", "status", "gender", "gender_confidence", "F9_07_COMP_DTK_COMP_ORG", "F9_07_COMP_DTK_EMPL_BEN", "F9_07_COMP_DTK_COMP_OTH")

for(i in 1:length(years)){
  link <- paste0("https://nccs-efile.s3.us-east-1.amazonaws.com/partvii/PARTVII-", years[i], "-PEOPLEPARSED.csv")
  temp <- fread(link, select = keep_cols_7) 
  colnames(temp) <- c("EIN", "year", "title", "first_name", "suffix", "status", "gender", "gender_confidence", "comp_main", "comp_benefit", "comp_other" )
  dat_7[[i]] <- temp
}

dat_all_7 <- rbindlist(dat_7)

dat_all_7 <-
  dat_all_7 %>% 
  filter(year <= 2013)


```


## Load in Part I data

```{r}
dat_1 <- vector(mode = "list", length = length(years))
keep_cols_1 <- c("ORG_EIN", "TAX_YEAR", "F9_01_ACT_GVRN_NUM_VOTE_MEMB", "F9_01_ACT_GVRN_NUM_VOTE_MEMB_IND")



for(i in 1:length(years)){
  link <- paste0("https://nccs-efile.s3.us-east-1.amazonaws.com/parsed/F9-P01-T00-SUMMARY-", years[i], ".csv")
  temp <- fread(link, select = keep_cols_1) 
  colnames(temp) <- c("EIN", "year", "num_vote_memb", "num_vote_memb_ind")
  dat_1[[i]] <- temp
}

dat_all_1 <- rbindlist(dat_1)

dat_all_1 <-
  dat_all_1 %>% 
  filter(year <= 2013)


```

## Basic Statistics 

### Number of board members per year From Part VII


```{r}
dat_all_7 %>% 
  group_by(EIN, year) %>% 
  summarise(num_members = n()) %>% 
  ungroup %>% 
  group_by(year) %>%
  summarise(avg_num_members = mean(num_members, na.rm=TRUE),
            sd_num_members = sd(num_members, na.rm= TRUE))


```

```{r}
dat_all_7 %>% 
  group_by(EIN, year) %>% 
  summarise(num_members = n()) %>% 
  ungroup %>%
  group_by(year) %>%
  ggplot(aes(x = num_members, y = after_stat(count),
             group = as.factor(year), color = as.factor(year) )) +
  geom_density(adjust = 2) +
  xlim(0, 50) +
  ggtitle("Number of Voting Members Listed on Part VII") +
  guides(color= guide_legend(title="Year"))+ 
  xlab("Number of Members") +
  ylab("Frequency")

```


### Number of board members per year From Part I


```{r}
#~34% of "Number of Members" are NA
mean(is.na(dat_all_1$num_vote_memb))


#~34% of "Number of Independent Members" are NA
mean(is.na(dat_all_1$num_vote_memb_ind))


dat_all_1 %>% 
  group_by(year) %>%
  summarise(avg_num_members = mean(num_vote_memb, na.rm=TRUE),
            sd_num_members = sd(num_vote_memb, na.rm= TRUE),
            avg_num_members_indep = mean(num_vote_memb_ind, na.rm=TRUE),
            sd_num_members_indep = sd(num_vote_memb_ind, na.rm= TRUE))

```

These numbers do not agree with the PART VII summary statistics. 


```{r}
dat_all_1 %>% 
  na.omit() %>%
  pivot_longer(!c(EIN, year), names_to = "type", values_to = "num") %>% 
  ggplot(aes(x = num, y = ..count..,
             color = as.factor(type))) +
  geom_density(adjust = 2) +
  facet_wrap(~year) +
  xlim(0, 50) +
  ggtitle("Number of Voting Members Listed on Part I") +
  guides(color=guide_legend(title="Type")) +
  xlab("Number of Members") +
  ylab("Frequency")


dat_all_1 %>% 
  na.omit() %>% 
  filter(num_vote_memb < 100, 
         num_vote_memb_ind < 100) %>%
  ggplot(aes(x = num_vote_memb, y = num_vote_memb_ind )) +
  geom_point(alpha = 0.05) +
  facet_wrap(~year) +
  ggtitle("Voting Members vs. Independent Voting Members") +
  xlab("Number of Voting Members") +
  ylab("Number of Independent Voting Members")

```




## Paid Members 

```{r}
#get number of members by org by year
by_ein_year_7 <- 
  dat_all_7 %>% 
  group_by(EIN, year) %>% 
  summarise(num_members = n()) 


#get sum of compensation for each person per year
num_zeros <- 
  dat_all_7 %>%
  rowwise() %>%
  mutate(comp_all = sum(comp_main, comp_benefit, comp_other, na.rm = TRUE))

#mark if they are zero or not
num_zeros$comp_all_0 <- num_zeros$comp_all == 0

dat_comp <-
  num_zeros %>% 
  filter(!comp_all_0) %>% 
  group_by(EIN, year) %>% 
  summarise(num_members_comp = n(),
            avg_comp = mean(comp_all))  %>% 
  left_join(by_ein_year_7[, c("EIN", "year", "num_members")] )
  

#number of compensated people per year
dat_comp %>% 
  filter(num_members < 101) %>%
  ggplot(aes(x = num_members_comp, y = avg_comp, color = num_members)) +
  geom_point(alpha = 0.25)+
  facet_wrap(~year) +
  xlim(0, 100) +
  ylim(0, 1000000) +
  ggtitle("Number of Compensated members Vs. Average Compensation") +
  xlab("Number of Compensated Members") +
  ylab("Average Total Compensation of Compensated Members")+
  scale_color_viridis()
  

#number of board member vs number of compensated board members

dat_comp %>% 
  filter(avg_comp < 1000001) %>% 
  ggplot(aes(x = num_members, y = num_members_comp, color = avg_comp)) +
  geom_point(alpha = 0.25)+
  facet_wrap(~year) +
  xlim(0, 100) +
  ylim(0, 100) +
  ggtitle("Number of Members Vs. Number of Compensated Members") +
  xlab("Number of Members") +
  ylab("Number of Compensated Members")+
  scale_color_viridis()



```


### Differnce in Number of Board Members From Year to Year

```{r}
### Change in Number of Board Members over the years
## only keeping the orgs that have data in every year
num_per_year <- 
  by_ein_year_7 %>% 
  pivot_wider(names_from = year, values_from = num_members) %>% 
  na.omit() %>% 
  relocate(EIN, all_of(as.character(years)))

#get 1 year difference, set year 1 at 0 for baseline
diff_by_year <- data.frame(EIN = num_per_year$EIN)
diff_by_year[, 2] <- 0
colnames(diff_by_year)[2] <- as.character(years[1])
for(i in 2:length(years)){
  diff_by_year[, i+1] <- num_per_year[, i+1] - num_per_year[, i]
}



diff_by_year %>% 
  pivot_longer(!EIN, names_to = "year", values_to = "diff") %>%
  ggplot(aes(y = diff, x = year)) +
  geom_boxplot(outlier.shape = NA) +
  ylim(-5, 5)  +
  ggtitle("Differnce in Number of Board Members From Year To Year") +
  ylab("difference from the previous year")


diff_by_year %>% 
  pivot_longer(!EIN, names_to = "year", values_to = "diff") %>%
  filter(year != 2009) %>% 
  ggplot(aes(x = diff)) +
  geom_histogram(bins = 20) +
  xlim(-10, 10) +
  facet_wrap(~year) +
  ggtitle("Differnce in Number of Board Members From Year To Year") +
  xlab("difference from the previous year")


```

### Percentage Differnce in Number of Board Members From Year to Year

```{r}
#Percentage Change by year 
perc_by_year <- data.frame(EIN = num_per_year$EIN)
perc_by_year[, 2] <- 0
colnames(perc_by_year)[2] <- as.character(years[1])
for(i in 2:length(years)){
  perc_by_year[, i+1] <- (num_per_year[, i+1] - num_per_year[, i]) /  num_per_year[, i] * 100
}
  
perc_by_year %>% 
  pivot_longer(!EIN, names_to = "year", values_to = "perc") %>%
  ggplot(aes(y = perc, x = year)) +
  geom_boxplot(outlier.shape = NA) +
  ylim(-25, 25)  +
  ggtitle("Percentage Differnce in Number of Board Members From Year To Year") +
  ylab("precentage difference from the previous year")

perc_by_year %>% 
  pivot_longer(!EIN, names_to = "year", values_to = "perc") %>%
  filter(year != 2009) %>% 
  ggplot(aes(x = perc)) +
  geom_histogram(bins = 20) +
  xlim(-50, 50) +
  facet_wrap(~year) +
  ggtitle("Percentage Difference in Number of Board Members From Year To Year") +
  xlab("percentage difference from the previous year")

  
```




## Gender Percentages 

```{r}
dat_all_7 %>% 
  group_by(EIN, year, gender) %>% 
  summarise(num_members = n() ) %>%
  ungroup %>%
  group_by(year, gender) %>%
  summarise(avg_num_members = mean(num_members, na.rm = TRUE)) %>% 
  arrange(desc(gender)) %>%
  ggplot(aes(x = year, y = avg_num_members, fill = fct_inorder(gender))) +
  geom_col(position="stack") +
  
  ggtitle("Average Board Members By Gender") +
  xlab("Year") +
  ylab("Average Number of Members")  +
  guides(fill=guide_legend(title="Gender"))
  

dat_all_7 %>% 
  group_by(EIN, year, gender) %>% 
  summarise(num_members = n() ) %>%
  ungroup %>%
  group_by(year, gender) %>%
  summarise(avg_num_members = mean(num_members, na.rm = TRUE)) %>% 
  arrange(desc(gender)) %>%
  ggplot(aes(x = year, y = avg_num_members, fill = fct_inorder(gender))) +
  geom_col(position="fill") +
  ggtitle("Percentage of Board Members By Gender") +
  xlab("Year") +
  ylab("Percentage of Members") +
  guides(fill=guide_legend(title="Gender")) 
```





